\section{INTRODUCTION}


\par In recent years, technological advances have led to a growing availability of massive amounts of continuous streaming data (i.e., data streams observing events) in many application domains such as social networks \cite{mathioudakis2010twittermonitor}, Internet of Things (IoT) \cite{miorandi2012internet}, and maritime surveillance \cite{patroumpas2015event}.  The ability to detect and predict the full matches of a pattern of interest (e.g., certain sequence of events), defined by a domain expert, is important for several operational decision making tasks.
\par An event stream is an unbounded collection of timely ordered data observations in the form of an attribute tuple that is composed of a value from finite event types along with other categorical and numerical attributes. In this work we deal with movement event streams. For instance, in the context of maritime surveillance, the event stream of a moving vessel consists of spatiotemporal and kinematic information along with the vessel's identification and its trajectory related events, based on the automatic identification system (AIS) \cite{ais} messages that are continuously sent by the vessel. Therefore, leveraging event patterns prediction over real-time streams of moving vessels is useful to alert maritime operation managers about suspicious activities (e.g., fast sailing vessels near ports, or illegal fishing) before they happen. However, processing real-time streaming data with low latency is challenging, since data streams are large and distributed in nature and continuously arrive at a high rate. 
% we describe the design and implementation of a system called% 
\par In this paper, we present the design and implementation of an online, distributed and scalable pattern prediction system over multiple, massive streams of events. More precisely, we consider event streams related to trajectories of moving objects (i.e., vessels). The proposed approach is based on a novel method that combines a distributed online prediction protocol \cite{dekel2012optimal,kamp2014communication} with an event forecasting method, based on Markov chains \cite{alevizos2017event}. It is implemented on top of the Big Data framework for stream processing Apache Flink \cite{Flink}. We evaluate our proposed system over real-word data streams of moving vessels, which are provided in the context of the datAcron project\footnote{\url{http://www.datacron-project.eu/}}.

\par The rest of the paper is organized as follows. We discuss the related work and used frameworks in Section ~\ref{sec:realred_work}. In Section ~\ref{sec:system}, we describe the problem of pattern prediction, our proposed approach, and the architecture of our system. The implementation details on top of Flink are presented in Section ~\ref{sec:impl} and the experimental results in Section ~\ref{sec:results}. We conclude in Section ~\ref{sec:concl}.
